---
title: "Updates to the R Documentatino System"
author: |
  Andrew Redd,
  R documentation Task Force
date: "UseR!2017"
output: slidy_presentation
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

## History

* UseR!2016
* July R Consortium Funding

* September-December 2016
* January 2017 - Now


## Design Principles

* Full class system for documentation
* Can document anything
  + stored in `attr(., ‘documentation’)` for traditional objects; 
    - functions, data.frames, etc.
* Stored as a meta object similar to class definitions for S4 and reference classes.
* Everything can be accessed or set from the `documentation` and `documentation<-` generics.
* Formats for input and output support decided
* Independent of but supports the package system


## Many in - Many Out

![Many in - Many out approach.](../Diagram1.svg)


## Current State

* `parsetools`
* `documentation`
* `documentation-md`
* `documentation-yaml

## `parsetools`

```{r example, echo=TRUE}
library(parsetools)
options(keep.source=TRUE)
pd <- get_parse_data(parse(text="x <- rnorm(100)"))
kable(pd[-(3:4)])
```


------
```{r, echo=TRUE}
get_pd_assign_value(pd)
get_pd_assign_variable(pd)
```


## `documentation`

```{r example-documentation}
library(documentation)
x <- new('Documentation'
        , title       = "Example documentation"
        , author      = person('Andrew', 'Redd')
        , description = "This is used to demonstrate the class" %\%
                        "system for the documentation package in R"
        , keywords    = "documentation"
        )
```

-----
```{r, echo=TRUE}
str(x)
```

-----
```{r, echo=TRUE}
toRd(x) %>% cat(sep='\n')
```

## Something a bit more interesting
```{r, message=FALSE, echo=TRUE}
hw <- function(){print("hello world")}
documentation(hw) <- function_documentation('hw', title = "the standard Hello world")
documentation(hw) %>% toRd %>% unlist %>% cat(sep='\n')
````


## Can your documentation do this?
```{r, echo=TRUE}
make_hello <- function(what = 'world'){
  documented(function(){
    print(paste("hello", what))
  }, name  = paste0("hello_", what)
   , title = sprintf("the dynamic hello_%s example", what)
   , description = "Prints 'hello" %\% what %\%"'to the console." %\n%
                   "Used to show the power of dynamic documentation."
   )
}
hello_world <- make_hello()
hello_belgium <- make_hello('Belgium')

hello_belgium()
```

-----
```{r}
documentation(hello_belgium) %>% toRd %>% unlist %>% cat(sep="\n")
```


## Test extraction

Example R code in "./R/hello.R"

```r
hello_belgium <- make_hello('Belgium')
if(FALSE){#@testing
  expect_output(hello_belgium(), "hello Belgium")
}

```

------

We run `extract_tests()` then ...

------

in "./tests/testthat/test-hello.R" we will have
```r
#! This file was automatically produced by documentation on 2017-07-05 11:00:00
#! changes will be overwritten.
context('tests extracted from file `./R/hello.R`')
test_that("hello_belgium", {#@testing
  expect_output(hello_belgium(), "hello Belgium")
})


```

## Something still to finish
**Relational Tags** 


```r
hw <- 
function( greet = "Hello" #< A greeting 
        , who   = "World" #< who to greet.
        ){
    #! A more complicated hello world
    print(paste(greet, who))
    #< Called for the side effect of printing, 
    #^ but returns the pasted arguments invisibly.
}
```


## State in summary 

* Documentation classes
    + done..*ish*
* parsetools 
    + functionally complete
* test extraction ... done
* Relational tags
    + infrastructure in place.
    + needs to be integrated into Roxygen.
* Input forms
    + lots of work still to do.
* Output forms
    + Rd inplace but could use improvement.
    + markdown lot to do
    + others not started

## how people can help and get involved.

* Make a github request for a feature.
* Code one of the missing functionalities
    + code it up
    + test the heck out of it
    + make a pull request showing

https://github.com/RDocTaskForce/

* Join the mailing list

https://lists.r-consortium.org/mailman/listinfo/rconsortium-wg-dtf

